### 2025-01-26 technical_doc/implementation.tex
finish converting some code to diagrams

\noindent The utilities module provide common tools that are used when working with the developed software. First, because the basis function used in this study is the Fourier basis, a large proportion of concern is the storage of Fourier basis coefficients. The Fourier series representation that is used in this study is the complex representation as show in \lccref{eq:fourier_series}. However, as previously mentioned, least squares support vector machines is constructed to work with real numbers. In order for LSSVR to learn the relationship between complex numbers, a representation of complex numbers in terms of real numbers is necessary. Since sample features and labels for LSSVR need to be flattened into one dimension which result in a matrix with rows of samples, the representation of complex numbers using real numbers can be done after features and labels have been flattened. An example of such a matrix of two samples with four features each can be seen in \lccref{eq:complex_matrix}. To represent the complex number features as real numbers, once can simply split a single complex number into two real numbers. When this is done to the entire matrix of complex numbers, the result is \lccref{eq:complex_matrix_real_rep}.
\begin{align}
  \begin{bmatrix}
    -1.03+0i & -0.52+0.39i & -0.54+0i & -0.52-0.39i \\
    -0.85+0i & -0.39+0.19i & 0.87+0i  & -0.39-0.19i
  \end{bmatrix}\label{eq:complex_matrix} \\
  \begin{bmatrix}
    -1.03 & 0 & -0.52 & 0.39 & -0.54 & 0 & -0.52 & -0.39 \\
    -0.85 & 0 & -0.39 & 0.19 & 0.87  & 0 & -0.39 & -0.19
  \end{bmatrix}\label{eq:complex_matrix_real_rep}
\end{align}
There are several ways to do this in PyTorch, such as extracting the real and imaginary element components into separate matrices and then interleaving the columns to create the real representation. However, a more performant version can be made by using built-in functions of the library that directly allows views of complex valued tensors as real tensors and vice versa. The \verb|view_as_real| and \verb|view_as_complex| functions allow the conversion between real and complex representations by representing complex valued tensors by adding a dimension of size two for each component of complex numbers. The resulting implementation can be seen in \lccref{fig:complex_matrix_conversion_implementation}.
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
def to_real_coeff(x: torch.Tensor) -> torch.Tensor:
    return torch.view_as_real(x).flatten(-2)

def to_complex_coeff(x: torch.Tensor) -> torch.Tensor:
  return torch.view_as_complex(
    x.reshape((*x.shape[:-1], -1, 2))
  )
  \end{lstlisting}
  \caption{Utility function to convert between complex matrices and the real representations.}\label{fig:complex_matrix_conversion_implementation}
\end{figure}

Another important utility used on inputs for the proposed SpectralSVR is a scaling function. This utility has been implemented in a variety of ways and using different tools. However, as the standard scaling is not provided by PyTorch, we have also implemented our own version. The scaling function accepts PyTorch tensors or tuples of tensors of real or complex elements. First, the scaling function is implemented as a class that is instantiated. The scaling function is then fitted onto the reference data to obtain the standard deviation and mean of each feature. This information is stored as properties of the class. Any complex tensors are converted into their real representations using the functions in \lccref{fig:complex_matrix_conversion_implementation}. Once the scaling function instance is fitted, any set of tensors with the same number of features (columns) and element type (complex or real) can be roughly transformed into a standard deviation of one and mean of zero for every feature. This is done by subtracting the fitted mean of each feature from the elements in the column and then dividing with the feature's standard deviation. The implementation also provides a method to retrieve a scaling function fitted on a subset of the original tuple of tensors. Serialization for saving to and loading from disk is also implemented. This is done simply by using the PyTorch save and load functions on the scaling function instance itself. For some use cases, an inverse scaling operation to the original standard deviation and mean is also essential. The implementation is simply the reverse of the transformation process where the elements are multiplied by the standard deviation and the mean is the added.

Another group of utility in the module is concerned with managing the number of modes a tensor of coefficients has. First, a function called \verb|resize_modes| accepts a tensor of coefficients and the target number of modes in each dimension. The first dimension is always assumed to represent the different samples. Because of this the target modes will only affect the dimensions after the first. The wave numbers are also assumed to be symmetric about zero such that the highest absolute values of the wave number is in the middle of each dimension. The wave number decreases symmetrically as you move towards the edges of the dimension and reaching the lowest absolute wave number at the edges. This is done to ensure compatibility with the way existing libraries such as PyTorch works with tensors of Fourier coefficients. Because of this, shrinking a tensor of coefficients along a dimension means removing the center of the tensor along that dimension. And expanding to a target number of modes that is larger means that the tensor is padded with zeros in the center along the desired dimension. In addition, the remaining coefficients may be rescaled in order to balance the effect of adding or removing coefficients. The implementation is presented in \lccrefs{fig:coefficient_shrink_impl,fig:coefficient_expand_impl}.

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
for dim, (target_mode, current_mode) in enumerate(
    zip(target_modes, current_modes), 1
):
    device = x.device
    start_range = torch.tensor(
        range((target_mode - 1) // 2 + 1), dtype=torch.int
    ).to(device=device)
    end_range = torch.tensor(
        range(current_mode - target_mode // 2,
          current_mode
        ),
        dtype=torch.int,
    ).to(device=device)

    x_resized = torch.concat(
        (
            x_resized.index_select(dim, start_range),
            x_resized.index_select(dim, end_range),
        ),
        dim,
    )
  \end{lstlisting}
  \caption{Example of shrinking a tensor of samples' coefficients to a target number of modes.}\label{fig:coefficient_shrink_impl}
\end{figure}
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
for dim, (target_mode, current_mode) in enumerate(
    zip(target_modes, current_modes), 1
):
    device = x.device
    start_range = torch.tensor(
        range((current_mode - 1) // 2 + 1), dtype=torch.int
    ).to(device=device)
    # make sure that end range is empty if
    # the coefficient is only size 1
    end_range = torch.tensor(
        range(current_mode // 2, current_mode)
        if current_mode > 1
        else range(0),
        dtype=torch.int,
    ).to(device=device)
    padding_size = target_mode - current_mode
    modes = list(x_resized.shape)
    modes[dim] = padding_size
    padding = torch.zeros(modes).to(x_resized)

    x_resized = torch.concat(
        (
            x_resized.index_select(dim, start_range),
            padding,
            x_resized.index_select(dim, end_range),
        ),
        dim,
    )
  \end{lstlisting}
  \caption{Example of expanding a tensor of samples' coefficients to a target number of modes.}\label{fig:coefficient_expand_impl}
\end{figure}
Next, the second way to manage the size of a tensor along a dimension is to interpolate the current values. The implementation uses simple linear interpolation. This is especially useful for situations such as coefficients that are themselves functions of time. This is a common application such as when solving the initial condition problem of a function in space and time. One might represent the function as a Fourier series with basis as functions of space and coefficients as functions of time. In this case, the coefficients for time values that are not available can be interpolated. The implementation of this is shown in \lccref{fig:coefficient_interpolation_impl}.
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
    index_floor = index_float.floor().to(torch.int)
    index_ceil = index_float.ceil().to(torch.int)
    x_ceil = x.index_select(dim, index_ceil)
    x_floor = x.index_select(dim, index_floor)
    # interpolate coefficients
    index_shape = [1 for _ in range(x_floor.ndim)]
    index_shape[1] = -1
    index_scaler = (
        ((index_float - index_floor) / (
          index_ceil - index_floor
        ))
        .reshape(index_shape)
        .nan_to_num()
    )
    # ynt + scaler * (ynt1 - ynt)
    # (1 - scaler) * ynt + scaler * ynt1
    x_interp = torch.lerp(x_floor, x_ceil, index_scaler.to(x))
  \end{lstlisting}
  \caption{Example of interpolating for indices (floating point indices) between the available ones (integer indices).}\label{fig:coefficient_interpolation_impl}
\end{figure}

Finally, this module also provides the types and reexports implementations of ordinary differential equation solvers from the torchdiffeq library \autocite{Chen_torchdiffeq_2021}. The implementation can be seen in \lccref{fig:solver_impl}.
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
RHSFuncType = Callable[[torch.Tensor, torch.Tensor],
  torch.Tensor
]
SolverSignatureType = Callable[[
        RHSFuncType, torch.Tensor, torch.Tensor,
    ],
    torch.Tensor
]
MixedRHSFuncType = Callable[[
        torch.Tensor, torch.Tensor, torch.Tensor,
    ],
    torch.Tensor,
]
MixedSolverSignatureType = Callable[
    [MixedRHSFuncType, torch.Tensor, torch.Tensor],
    torch.Tensor,
]

def euler_solver(
    rhs_func: RHSFuncType,
    y0: torch.Tensor,
    t: torch.Tensor,
):
    solution = torch.zeros((len(t), *y0.shape)).to(y0)

    j = 1
    solution[j - 1] = y0
    for t0, t1 in zip(t[:-1], t[1:]):
        dt = t1 - t0
        y = solution[j - 1]
        solution[j] = y + dt * rhs_func(t0, y)
        assert (
            solution.isnan().sum() == 0
        ), f"solver encountered nan at timestep {j} (t={t0})"
        j = j + 1
    return solution

implicit_adams_solver: SolverSignatureType = partial(
    odeint, method="implicit_adams", options={"max_iters": 4}
)  # type: ignore

lsoda_solver: SolverSignatureType = partial(
    odeint, method="scipy_solver", options={"solver": "LSODA"}
)  # type: ignore
  \end{lstlisting}
  \caption{Implementation of ODE solvers and the types.}\label{fig:solver_impl}
\end{figure}

@@

\noindent This module implements two classes. The first is the base Basis class. The second is a subclass implementing the Fourier basis specifically. These classes provide the functionality needed to have an easier time when working with basis functions and their coefficients. First, the core function of these classes is to provide a way to transition between the spectral domain and the physical domain. This is implemented in the base class by storing the coefficients of sample functions as a class property. The coefficients are assumed to be at least of one function.

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    @staticmethod
    def transform(
        f: torch.Tensor,
        res: TransformResType | None = None,
        periodic: bool = True,
        periods: PeriodsInputType = None,
        allow_fft: bool = True,
    ) -> torch.Tensor:
        if not torch.is_complex(f):
            f = f * (1 + 0j)
        res = transformResType_to_tuple(res, tuple(f.shape[1:]))
        periods = periodsInputType_to_tuple(periods, f.shape[1:])
        # perform 1d transform over every dimension
        F = f
        for cdim in range(1, ndims):
            F = FourierBasis._ndim_transform(
                F,
                dim=cdim,
                func="forward",
                res=res[cdim - 1],
                periodic=periodic,
                period=periods[cdim - 1],
                allow_fft=allow_fft,
            )

        return F
...
  \end{lstlisting}
  \caption{Implementation of forward Fourier transform.}\label{fig:fourier_transform_impl}
\end{figure}

The transitions between the coefficients and functions values are implemented as a function to transform function values to coefficients and an inverse transform function to compute function values from coefficients. The base class enforces this by defining abstract methods which any subclass will need to implement. The FourierBasis subclass then implements the transform and inverse transform functions specific for transitioning between function values and Fourier basis coefficients. As before, the implementation assumes that the first dimension indexes the samples and the rest represent the coefficient wave numbers. This means that the implementation must allow for multidimensional functions. For the Fourier transform and inverse transform, the multidimensional version is relatively simple to implement. The multidimensional Fourier transform essentially performs the one dimensional Fourier transform along one dimension. And then, using the result to perform the one dimensional Fourier transform along the next dimension. This process is repeated until the last dimension. The implementation of the forward transform is presented in \lccref{fig:fourier_transform_impl}. The inverse transform simply changes the \verb|func| parameter to \enquote{inverse}.

The Fourier transform along each dimension is done by flattening every other dimension into the sample dimension. This essentially means that the function values are \enquote{sliced} along the dimension currently being transformed, and that each \enquote{slice} is its own \enquote{sample}. Mathematically, this is simply due to the matrix multiplication of basis functions along one dimension with every other dimension. The implementation of this is shown in \lccref{fig:ndim_fourier_transform_impl}.

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    @staticmethod
    def _ndim_transform(
        f: torch.Tensor,
        dim: int,
        func: Literal["forward", "inverse"],
        res: slice,
        periodic: bool,
        period: float,
        allow_fft: bool,
    ) -> torch.Tensor:
        # flatten so that each extra dimension is
        # treated as a separate "sample"
        # move dimension to transform to the end
        # so that it can stay intact after f is flatened
        f_transposed = f.moveaxis(dim, -1)
        # flatten so that the last dimension is intact
        f_flatened = f_transposed.flatten(0, -2)

        F_flattened = FourierBasis._raw_transform(
            f_flatened,
            func=func,
            res=res,
            periodic=periodic,
            period=period,
            allow_fft=allow_fft,
        )
        # unflatten so that the correct shape is returned
        F_transposed = F_flattened.reshape(
          (*f_transposed.shape[:-1], res.step)
        )
        F = F_transposed.moveaxis(-1, dim)

        return F
...
  \end{lstlisting}
  \caption{Implementation of n-dimensional Fourier transform for a specific dimension.}\label{fig:ndim_fourier_transform_impl}
\end{figure}

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    @staticmethod
    def _raw_transform(
        f: torch.Tensor,
        func: Literal["forward", "inverse"],
        res: slice,
        periodic: bool,
        period: float,
        allow_fft: bool,
    ) -> torch.Tensor:
        match func:
            case "forward":
                sign = -1
            case "inverse":
                sign = 1
        mode = f.shape[1]
        domain_starts_at_0 = res.start == 0
        domain_end_equal_to_period = res.stop == period
        can_use_fft = (
            domain_starts_at_0
            and domain_end_equal_to_period
            and periodic
            and allow_fft
        )
        if can_use_fft:
            if func == "forward":
                F = torch.fft.fft(
                  f, dim=1, n=res.step, norm="backward"
                )
            elif func == "inverse":
                F = torch.fft.ifft(
                  f, dim=1, n=res.step, norm="forward"
                )
        else:
            if periodic:
                n = res.start + torch.arange(res.step).to(f)
                n = n / res.step * period
            else:
                n = torch.linspace(
                    res.start, res.stop, res.step
                ).to(f)
            e = FourierBasis.fn(
                n.view(-1, 1),
                mode,
                periods=period,
                constant=sign * 2j * torch.pi,
            )

            F = torch.mm(f, e.T)
        return F
...
  \end{lstlisting}
  \caption{Implementation of the one dimensional Fourier Transform.}\label{fig:1d_fourier_transform_impl}
\end{figure}

The one dimensional Fourier transform itself is implemented as a raw transform function. This function accepts input of a matrix of values to transform, a parameter to indicate the type of transformation (inverse transform or forward transform), the evaluation boundary and number of grid points, whether the evaluation should assume the function is periodic, the size of the function domain, and if the Fast Fourier Transform is allowed to be used. The evaluation boundaries, actual function domain size, whether the function can be assumed to be periodic, and if FFTs is allowed to be used together determine the Discrete Fourier Transform algorithm that is used. The FFT algorithm is used if the following are all true: the evaluation starts at zero and ends with the same value as the period, the function can be assumed to be periodic, and FFT is allowed. Otherwise, the transform is computed naively using \lccrefs{eq:discrete_fourier_transform,eq:inverse_discrete_fourier_transform}. The naive approach allows for evaluating function values of coefficients at various grid sizes and evaluation boundaries. This is useful for being able to evaluate at resolutions other than that of the original discretized function or coefficients. One use for this is plotting the function at different parts or resolutions. The relationship with the physical domain means that the boundaries of the domain is important. This is because the function values that will be used may only be valid within certain boundaries. Because of this, the information of physical domain bounds must also be stored. This is done by storing the span of the function in each dimension. The implementation we have gone with doesn't store any other information than the span in each dimension. Because of this, the information would need to be stored outside the Basis class instance and any outputs or operations with the classes will need to take this into account. The implementation of the one dimensional Fourier transform is presented in \lccref{fig:1d_fourier_transform_impl}.

The secondary set of functionality provided by the basis classes are mathematical operators on the functions the coefficients represent. There are four operators implemented, which are the addition, subtraction, derivative, and antiderivative operators. These four operations can be further categorized into arithmetic and calculus operations. The arithmetic operations, which are addition and subtraction, are implemented by following how the actual mathematical operations would be carried out on two functions which are Fourier series as shown in \lccref{eq:fourier_series_addition}. For subtraction, the plus sign is simply replaced with the minus sign.
\begin{align}
  \sum_{k} \hat{u}_k e^{2\pi ikx} + \sum_{k} \hat{f}_k e^{2\pi ikx} =
  \sum_{k} \left(\hat{u}_k + \hat{f}_k\right) e^{2\pi ikx} \label{eq:fourier_series_addition}
\end{align}
The implementation leverages the operator overloading in Python classes. In addition, the overload should be done to the \verb|__add__| function. This function adds the current instance to another object which is the other basis instance for our use case. We also want the operation to return an independent instance without any references to the previous instances. To do this, all properties are copied into a new instance. This is then used to perform the arithmetic operation. The implementation in \lccref{fig:basis_addition_impl} demonstrates the addition operation. For subtraction, a similar implementation is done using the \verb|__sub__| function and switching the plus sign with the minus sign.

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    def __add__(self, other: Self):
        if isinstance(other, self.__class__):
            if other.coeff is None:
                return self.copy()
            elif self.coeff is None:
                return other.copy()
            else:
                result = self.resize_modes(other)
                result.coeff = result.coeff + other.coeff
                return result
        else:
            raise TypeError(
                f"unsupported operand type(s) for +: '{
                    self.__class__
                }' and '{type(other)}'"
            )
...
  \end{lstlisting}
  \caption{Implementation of basis addition function.}\label{fig:basis_addition_impl}
\end{figure}

The calculus operations for Fourier basis take advantage of the properties of the Fourier series. As discussed in \lccref{sec:fourier_discussion}, derivatives and integration become multiplication and division in the spectral space of Fourier series. Our implementation takes advantage of this as shown in \lccref{fig:basis_grad_impl}. Similar to how subtraction is just a modification of the addition operation, the integral or antiderivative operator is also implemented simply by changing the operation with the term multiplier from multiplication to division.

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    def grad(self, dim: int = 0, ord: int = 1) -> Self:
        copy = self.copy()
        if dim == 0 and self.time_dependent:
            # time dependent use finite differences
            dt = self.periods[0] / (self.time_size - 1)
            coeff = copy.coeff
            for o in range(ord):
                coeff = torch.gradient(coeff, spacing=dt, dim=1)[0]
            copy.coeff = coeff
        else:
            if self.time_dependent:
                # disregard time dimension
                dim = dim - 1
            k = copy.wave_number(copy.modes[dim])
            multiplier_dims = [1 for _ in range(copy.ndim)]
            multiplier_dims[dim] = copy.modes[dim]
            if self.time_dependent:
                multiplier_dims = (1, *multiplier_dims)
            multiplier = (
                2
                * torch.pi
                * 1j
                * k.reshape(multiplier_dims).to(copy.coeff)
                / self.periods[dim]
            )
            multiplier = multiplier.pow(ord)
            coeff = copy.coeff.mul(multiplier)
            coeff[:, ..., 0] = torch.tensor(0 + 0j)
            copy.coeff = coeff
        return copy
...
  \end{lstlisting}
  \caption{Implementation of basis grad function.}\label{fig:basis_grad_impl}
\end{figure}

The last group of functionality implemented in the Basis and FourierBasis classes provides convenience and often used procedures when working with basis functions. First, random coefficient generation for the Fourier basis is implemented. This is done as discussed in step 3 of the data generation process discussed in \lccref{sec:data_generation}. The implemented function shown in \lccref{fig:generate_coeff_impl} is then wrapped in another function that creates a new FourierBasis instance with the generated coefficients.

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    @classmethod
    def generate_coeff(
        cls,
        n: int,
        modes: int | tuple[int, ...],
        generator: torch.Generator | None = None,
        random_func: Callable[..., torch.Tensor] = torch.randn,
        complex_funcs: bool = False,
        scale: bool = True,
    ) -> torch.Tensor:
        if isinstance(modes, int):
            modes = (modes,)
        n_modes = len(modes)
        assert n_modes > 0, "modes should have at least one element"
        random_func = partial(random_func, generator=generator)
        if complex_funcs:
            coeff = random_func((n, *modes), dtype=torch.complex64)
        else:
            coeff = random_func((n, *modes), dtype=torch.complex64)

            vals = cls.inv_transform(coeff)
            coeff = cls.transform(vals.real + 0j)
        if scale:
            scaler = torch.tensor(modes).sum() * 0.2
            coeff = coeff.mul(scaler)
        return coeff
...
  \end{lstlisting}
  \caption{Implementation of Fourier Basis coefficient generation function.}\label{fig:generate_coeff_impl}
\end{figure}

@@

The optimization function that fits the model to the training data solves \lccref{eq:lssvr_solution}. This function is shown in \lccref{fig:lssvr_fit_function_impl}. The first half of the function sets up the left and right hand side matrices. The left hand side matrix is constructed in part with the kernel matrix. The kernel function used is batched to reduce the peak memory footprint. These matrices are then used in the second half to compute the solution using the least squares solver function from PyTorch called lstq. Finally, the result is split into the biases and Lagrange multipliers which are then returned as a tuple. The optimizing function is wrapped in a more user-friendly function which allows the use of NumPy arrays in addition to the PyTorch Tensor used in the optimizing function. The wrapper function is named fit. It conforms the inputs provided by the user into the expected types and matrix shapes of rows for samples and columns for input features or output targets. The processed inputs and outputs are also stored as support vectors for use during prediction. The fit function returns the class instance once training is done. This is for method chaining purposes for ease of use when using the fit function in a chain with other functions we will discuss next, such as the prediction function.

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    def _optimize_parameters(
      self,
      X: torch.Tensor,
      y_values: torch.Tensor,
    ):

        A = torch.empty(
          (X.shape[0] + 1,) * 2,
          device=self.device,
          dtype=self.dtype,
        )
        A[1:, 1:] = self._batched_K(X, X)
        A[1:, 1:].diagonal().copy_(
            A[1:, 1:].diagonal()
            + torch.ones(
                (A[1:, 1:].shape[0],),
                device=self.device,
                dtype=self.dtype,
            ).to()
            / self.C
        )
        A[0, 0] = 0
        A[0, 1:] = 1
        A[1:, 0] = 1
        shape = np.array(y_values.shape)
        shape[0] += 1
        B = torch.empty(
          list(shape), device=self.device, dtype=self.dtype
        )
        B[0] = 0
        B[1:] = y_values

>>>>

\noindent The utilities module provide common tools that are used when working with the developed software. First, because the basis function used in this study is the Fourier basis, a large proportion of concern is the storage of Fourier basis coefficients. However, as previously mentioned, least-squares support vector machines is constructed to work with real numbers. In order for LSSVR to learn the relationship between complex numbers, a representation of complex numbers in terms of real numbers is necessary. Since sample features and labels for LSSVR need to be flattened into one dimension which result in a matrix with rows of samples, the representation of complex numbers using real numbers can be done after features and labels have been flattened. An example of such a matrix of two samples with four features each can be seen in \lccref{eq:complex_matrix}. To represent the complex number features as real numbers, once can simply split a single complex number into two real numbers. When this is done to the entire matrix of complex numbers, the result is \lccref{eq:complex_matrix_real_rep}.
\begin{align}
    \begin{bmatrix}
        -1.03+0i & -0.52+0.39i & -0.54+0i & -0.52-0.39i \\
        -0.85+0i & -0.39+0.19i & 0.87+0i  & -0.39-0.19i
    \end{bmatrix}\label{eq:complex_matrix} \\
    \begin{bmatrix}
        -1.03 & 0 & -0.52 & 0.39 & -0.54 & 0 & -0.52 & -0.39 \\
        -0.85 & 0 & -0.39 & 0.19 & 0.87  & 0 & -0.39 & -0.19
    \end{bmatrix}\label{eq:complex_matrix_real_rep}
\end{align}
There are several ways to do this in PyTorch, such as extracting the real and imaginary element components into separate matrices and then interleaving the columns to create the real representation. However, a more performant version can be made by using built-in functions of the library that directly allows views of complex valued tensors as real tensors and vice versa. The \verb|view_as_real| and \verb|view_as_complex| functions allow the conversion between real and complex representations by representing complex valued tensors by adding a dimension of size two for each component of complex numbers. The resulting implementation can be seen in \lccref{fig:complex_matrix_format}.
\begin{figure}[H]
    \centering
    \tikzfig{figures/complex_number_real_format}
    \caption{Utility function to convert between complex matrices and the real representations.}\label{fig:complex_matrix_format}
\end{figure}

Another important utility used on inputs for the proposed SpectralSVR is a scaling function. This utility has been implemented in a variety of ways and using different tools. However, as the standard scaling is not provided by PyTorch, we have also implemented our own version. The scaling function accepts PyTorch tensors or tuples of tensors of real or complex elements. First, the scaling function is implemented as a class that is instantiated. The scaling function is then fitted onto the reference data to obtain the standard deviation and mean of each feature. This information is stored as properties of the class. Any complex tensors are converted into their real representations. Once the scaling function instance is fitted, any set of tensors with the same number of features (columns) and element type (complex or real) can be roughly transformed into a standard deviation of one and mean of zero for every feature. This is done by subtracting the fitted mean of each feature from the elements in the column and then dividing with the feature's standard deviation. The implementation also provides a method to retrieve a scaling function fitted on a subset of the original tuple of tensors. Serialization for saving to and loading from disk is also implemented. This is done simply by using the PyTorch save and load functions on the scaling function instance itself. For some use cases, an inverse scaling operation to the original standard deviation and mean is also essential. The implementation is simply the reverse of the transformation process where the elements are multiplied by the standard deviation and the mean is the added.

Another group of utility in the module is concerned with managing the number of modes a tensor of coefficients has. First, a function called \verb|resize_modes| accepts a tensor of coefficients and the target number of modes in each dimension. The first dimension is always assumed to represent the different samples. Because of this the target modes will only affect the dimensions after the first. The wave numbers are also assumed to be symmetric about zero such that the highest absolute values of the wave number is in the middle of each dimension. The wave number decreases symmetrically as you move towards the edges of the dimension and reaching the lowest absolute wave number at the edges. This is done to ensure compatibility with the way existing libraries such as PyTorch works with tensors of Fourier coefficients. Because of this, shrinking a tensor of coefficients along a dimension means removing the center of the tensor along that dimension. And expanding to a target number of modes that is larger means that the tensor is padded with zeros in the center along the desired dimension. In addition, the remaining coefficients may be rescaled in order to balance the effect of adding or removing coefficients. The implementation is presented in \lccrefs{fig:resize_modes_process}.

\begin{figure}[H]
    \begin{subfigure}{0.45\textwidth}
    \centering
        \tikzfig{figures/reduce_modes}
        \caption{}\label{fig:reduce_modes}
    \end{subfigure}
    \begin{subfigure}{0.45\textwidth}
    \centering
        \tikzfig{figures/expand_modes}
        \caption{}\label{fig:expand_modes}
    \end{subfigure}
    \caption{The two ways to resize coefficient tensors: (\subref{fig:reduce_modes}) Reduce coefficient modes, (\subref{fig:expand_modes}) Expand coefficient modes.}\label{fig:resize_modes_process}
\end{figure}

Next, the second way to manage the size of a tensor along a dimension is to interpolate the current values. The implementation uses simple linear interpolation. This is especially useful for situations such as coefficients that are themselves functions of time. This is a common application such as when solving the initial condition problem of a function in space and time. One might represent the function as a Fourier series with basis as functions of space and coefficients as functions of time. In this case, the coefficients for time values that are not available can be interpolated. The visualization of this is shown in \lccref{fig:linear_interpolation}.
\begin{figure}[H]
    \centering
    \tikzfig{figures/linear_interpolation}
    \caption{Linear interpolation process}\label{fig:linear_interpolation}
\end{figure}

Finally, this module also provides the types and reexports implementations of ordinary differential equation solvers from the torchdiffeq library \autocite{Chen_torchdiffeq_2021}. This portion of the module provide a uniform interface to use for the rest of the library. This is needed because of the extensibility requirement and the possibility of using other numerical method implementations.

@@

\noindent This module implements two classes. The first is the base Basis class. The second is a subclass implementing the Fourier basis specifically. These classes provide the functionality needed to have an easier time when working with basis functions and their coefficients. First, the core function of these classes is to provide a way to transition between the spectral domain and the physical domain. This is implemented in the base class by storing the coefficients of sample functions as a class property. The coefficients are assumed to be at least of one function.
\begin{figure}[H]
    \centering
    \tikzfig{figures/fourier_transform_process}
    \caption{Fourier transforms for multiple dimensions.}\label{fig:fourier_transform_multidim}
\end{figure}

The transitions between the coefficients and functions values are implemented as a function to transform function values to coefficients and an inverse transform function to compute function values from coefficients. The base class enforces this by defining abstract methods which any subclass will need to implement. The FourierBasis subclass then implements the transform and inverse transform functions specific for transitioning between function values and Fourier basis coefficients. As before, the implementation assumes that the first dimension indexes the samples and the rest represent the coefficient wave numbers. This means that the implementation must allow for multidimensional functions. For the Fourier transform and inverse transform, the multidimensional version is relatively simple to implement. The multidimensional Fourier transform essentially performs the one dimensional Fourier transform along one dimension. And then, using the result to perform the one dimensional Fourier transform along the next dimension. This process is repeated until the last dimension. The visualization of the forward transform for multiple dimensions is presented in \lccref{fig:fourier_transform_multidim}. The inverse transform simply changes the \verb|func| parameter to \enquote{inverse}.

The Fourier transform along each dimension is done by flattening every other dimension into the sample dimension. This essentially means that the function values are \enquote{sliced} along the dimension currently being transformed, and that each \enquote{slice} is its own \enquote{sample}. Mathematically, this is simply due to the matrix multiplication of basis functions along one dimension with every other dimension. The visualization of this process is shown in \lccref{fig:fourier_transform_along_one_dim}.
\begin{figure}[H]
    \centering
    \tikzfig{figures/fourier_transform_along_one_dim}
    \caption{The application of the one dimension Fourier transform on multiple dimensions.}\label{fig:fourier_transform_along_one_dim}
\end{figure}
\begin{figure}[H]
    \centering
    \tikzfig{figures/fourier_transform_one_dim}
    \caption{The application of the one dimension Fourier transform on multiple samples.}\label{fig:fourier_transform_one_dim}
\end{figure}

The one dimensional Fourier transform itself is implemented as a raw transform function. This function accepts input of a matrix of values to transform, a parameter to indicate the type of transformation (inverse transform or forward transform), the evaluation boundary and number of grid points, whether the evaluation should assume the function is periodic, the size of the function domain, and if the Fast Fourier Transform is allowed to be used. The evaluation boundaries, actual function domain size, whether the function can be assumed to be periodic, and if FFTs is allowed to be used together determine the Discrete Fourier Transform algorithm that is used. The FFT algorithm is used if the following are all true: the evaluation starts at zero and ends with the same value as the period, the function can be assumed to be periodic, and FFT is allowed. Otherwise, the transform is computed naively using the Discrete Fourier Transform (DFT) and Inverse Discrete Fourier Transform (IDFT). The naive approach allows for evaluating function values of coefficients at various grid sizes and evaluation boundaries. This is useful for being able to evaluate at resolutions other than that of the original discretized function or coefficients. One use for this is plotting the function at different parts or resolutions. The relationship with the physical domain means that the boundaries of the domain is important. This is because the function values that will be used may only be valid within certain boundaries. Because of this, the information of physical domain bounds must also be stored. This is done by storing the span of the function in each dimension. The implementation we have gone with doesn't store any other information than the span in each dimension. Because of this, the information would need to be stored outside the Basis class instance and any outputs or operations with the classes will need to take this into account. The implementation of the one dimensional Fourier transform is presented in \lccref{fig:fourier_transform_one_dim}.

The secondary set of functionality provided by the basis classes are mathematical operators on the functions the coefficients represent. There are four operators implemented, which are the addition, subtraction, derivative, and antiderivative operators. These four operations can be further categorized into arithmetic and calculus operations. The arithmetic operations, which are addition and subtraction, are implemented by following how the actual mathematical operations would be carried out on two functions which are Fourier series as shown in \lccref{eq:fourier_series_addition}. For subtraction, the plus sign is simply replaced with the minus sign.
\begin{align}
    \sum_{k} \hat{u}_k e^{2\pi ikx} + \sum_{k} \hat{f}_k e^{2\pi ikx} =
    \sum_{k} \left(\hat{u}_k + \hat{f}_k\right) e^{2\pi ikx} \label{eq:fourier_series_addition}
\end{align}
The implementation leverages the operator overloading in Python classes. In addition, the overload should be done to the \verb|__add__| function. This function adds the current instance to another object which is the other basis instance for our use case. We also want the operation to return an independent instance without any references to the previous instances. To do this, all properties are copied into a new instance. This is then used to perform the arithmetic operation. For subtraction, a similar implementation is done using the \verb|__sub__| function and switching the plus sign with the minus sign.

The calculus operations for Fourier basis take advantage of the properties of the Fourier series. Derivatives and integration become multiplication and division in the spectral space of Fourier series. Similar to how subtraction is just a modification of the addition operation, the integral or antiderivative operator is also implemented simply by changing the operation with the term multiplier from multiplication to division.

The last group of functionality implemented in the Basis and FourierBasis classes provides convenience and often used procedures when working with basis functions. First, random coefficient generation for the Fourier basis is implemented. The function is then wrapped in another function that creates a new FourierBasis instance with the generated coefficients.

@@

The optimization function that fits the model to the training data solves the linear optimization problem of the LSSVR\@. This process is shown in \lccref{fig:lssvr_optimization}. The first half of the function sets up the left and right hand side matrices. The left hand side matrix is constructed in part with the kernel matrix. The kernel function used is batched to reduce the peak memory footprint. These matrices are then used in the second half to compute the solution using the least squares solver function from PyTorch called lstq. Finally, the result is split into the biases and Lagrange multipliers which are then returned as a tuple. The optimizing function is wrapped in a more user-friendly function which allows the use of NumPy arrays in addition to the PyTorch Tensor used in the optimizing function. The wrapper function is named fit. It conforms the inputs provided by the user into the expected types and matrix shapes of rows for samples and columns for input features or output targets. The processed inputs and outputs are also stored as support vectors for use during prediction. The fit function returns the class instance once training is done. This is for method chaining purposes for ease of use when using the fit function in a chain with other functions we will discuss next, such as the prediction function.
\begin{figure}[H]
    \centering
    \tikzfig{figures/lssvr_optimization}
    \caption{Optimization process of LSSVR}\label{fig:lssvr_optimization}
\end{figure}

The learned multipliers and bias parameters are then used for predicting outputs from unseen sample features. The prediction process first constructs both matrices on the left-hand side of the equation. This is done by computing the kernel matrix between the prediction features and the training features, otherwise known as support vectors. The prediction function therefore is very straight forward with an input of unseen features and the learned parameters. The function output is simply the predicted outputs. The implementation of this function is shown in \lccref{fig:lssvr_optimization}.
\begin{figure}[H]
    \centering
    \tikzfig{figures/lssvr_prediction}
    \caption{Prediction process of LSSVR}\label{fig:lssvr_prediction}
\end{figure}
